Analysis of cloud enhancement events in a 30-year record of global solar irradiance at Thessaloniki, Greece

### 2024-09-08 article/article.tex
Commit 2024-09-08 12:14

In this study, we investigate the characteristics of global horizontal
irradiance enhancement events induced by clouds over Thessaloniki for
the period 1994 -- 2023 using data recorded every one minute. We
identified the cloud enhancement (CE) events by creating an appropriate
cloud-free irradiance reference using a radiative transfer model and
aerosol optical depth data from a collocated Cimel sun photometer and a
Brewer spectrophotometer. We found a trend in CE events of
\(+111\pm 27\,\text{cases}/\text{year}\), and a trend in the
corresponding irradiation of \(+325.2\pm 88\,\text{kJ}/\text{year}\).
The peak of the CE events was observed during May and June. CE events
with duration longer than 10\nobreakspace{}min are very rare
(\(<8\,\%\)), with exceptions lasting over an hour and up to 140
minutes. Finally, we have detected enhancements above the total solar
irradiance at the top of the atmosphere for the same solar zenith angle
of up to \(400\,\text{W}/\text{m}^{2}\), with the \(75\,\%\) of the
cases below \(200\,\text{W}/\text{m}^{2}\). Most of these extreme events
occur in spring -- early summer, with a secondary peak in autumn.

>>>>

In this study, we investigate the characteristics of global horizontal
irradiance enhancement events induced by clouds over Thessaloniki for
the period 1994 -- 2023 using data recorded every one minute. We
identified the cloud enhancement (CE) events by creating an appropriate
cloud-free irradiance reference using a radiative transfer model and
aerosol optical depth data from a collocated Cimel sun photometer and a
Brewer spectrophotometer. We found a trend in CE events of
\(+111\pm 27\,\text{cases}/\text{year}\), and a trend in the
corresponding irradiation of \(+325.2\pm 88\,\text{kJ}/\text{year}\).
The peak of the CE events was observed during May and June. CE events
with duration longer than 10\nobreakspace{}min are very rare
(\(<8\,\%\)), with exceptions lasting over an hour and up to 140
minutes. Finally, we have detected enhancements above the total solar
irradiance at the top of the atmosphere for the same solar zenith angle
of up to \(209\,\text{W}/\text{m}^{2}\), with the \(75\,\%\) of the
cases below \(40\,\text{W}/\text{m}^{2}\). Most of these extreme events
occur in spring -- early summer, with a secondary peak in autumn.
